---
title: "Roman Emperors"
author: "Bela Szlovicsak"
date: "2022-12-23"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidytuesdayR)
library(tidyverse)
library(broom)
library(easystats)

```

## Exploratory Analysis

Reading the data

```{r data read }
emperors <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-08-13/emperors.csv")

```


Looking at basic values, and examining the variables.

```{r}
str(emperors)

summary(emperors)

```

The dates are good if we need difference, because BCE dates are one less than in reality
Next I make two variables, one of reign length, another of age at death.

```{r}
emperors_dur <- emperors %>% 
  mutate(
    age = as.numeric(death-birth)/365,
    reign = as.numeric(reign_end - reign_start)/365
  ) %>% 
  filter(age > 0)

```

#Exploring possible relationships

 A relationship between age and length of reign?
```{r}
ggplot(emperors_dur, aes(age, reign)) +
  geom_point()

```

Examining reign vs. cause of death 
```{r}
ggplot(emperors_dur, aes(cause)) +
  geom_bar()


ggplot(emperors_dur, aes(cause, reign)) +
  geom_boxplot() 
```


Examining reign vs. mode of rising to power 

```{r}
ggplot(emperors_dur, aes(rise)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90))


ggplot(emperors_dur, aes(rise, reign)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90))
```

Examining cause of death vs. mode of rising to power
```{r}
emperors_dur %>% 
  count(rise, cause) %>% 
  ggplot( aes(rise, cause, fill = n)) +
  geom_tile() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_fill_viridis_c() +
  theme_minimal()



```

This last relationship seems worthy of examination, this could be done with a Fisher test

```{r}
emperors_dur %>% 
  count(rise, cause) %>% 
  pivot_wider(names_from = cause, values_from = n) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("rise") %>% 
  as.matrix() %>% 
  fisher.test(workspace = 20000000)

```

Because of the very low counts, no real relationship can be estabilished



Examining age vs. cause of death
```{r}
 



ggplot(emperors_dur, aes(cause, age)) +
  geom_boxplot() 

```


Examining age vs. mode of rising to power 

```{r}


ggplot(emperors_dur, aes(rise, age)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90))

```

Both rise and cause seem to have an effect on age, this can be checked with a linear model
  
```{r}
model1 <- lm(age ~ rise + cause +0, emperors_dur) 

tidy(model1 ,conf.int = TRUE, conf.level = 0.95)

```

Almost all confidence intervals for the causes of death contain 0, therefore this variable is not really adding anytting to this model, I'll make a new one without including it.


```{r}
model2 <- lm(age ~ rise +0, emperors_dur)
  
tidy(model2, conf.int = TRUE, conf.level = 0.95)

plot(model2)

```

Most technical requirements are okay, but the data is not exactly heteroscadistic, therefore it's not reliable.

